---
title: "Exploratory Data Analysis"
subtitle: "Pro-environmental Project"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



```{r libraries, include = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
library(rio) # for importing data
library(tidyverse) # for tidyverse tools
library(psych) # for descriptive statistics
library(Hmisc) # for histogram matrix
library(readr) # for exporting csv files
library(corrplot) # for a correlation matrix plot
library(naniar) # for visualizing missingness & MCAR test
library(cowplot) # for plotting multiple visualizations
library(GGally) # for scatterplot matrices
```


# Importing Data {.tabset .tabset-pill}

## Import Data
```{r}
data <- import("proenv_clean_data.csv")
```

## Check Measure Types

```{r inspect measure types, results = 'hide'}
str(data, list.len = ncol(data))
```

## Transform Measure Types

```{r transform measure types, warning=FALSE}
data <- data %>%
  mutate(id = as.factor(id),
         consumer_behaviors = as.factor(consumer_behaviors),
         framing_condition = as.factor(framing_condition),
         norm_condition = as.factor(norm_condition),
         Gender = as.factor(Gender),
         Class_Lvl = as.factor(Class_Lvl),
         Employment = as.factor(Employment),
         Parents_Education = as.factor(Parents_Education),
         Pol_Ornt = factor(Pol_Ornt, labels = c("Very liberal", "Somewhat liberal", "Slightly liberal", "Neither liberal nor conservative", "Slightly conservative", "Somewhat conservative", "Very conservative", "Other")),
         Ethnicity = as.factor(Ethnicity))
```

## Subset Key Variables
```{r}
data <- data %>%
  dplyr::select(id, framing_condition, norm_condition, biospheric, egoistic, clothing_interest, ingroup_identification, consumer_intentions, consumer_behaviors, Age, Gender)
```


# Descriptive Tables {.tabset .tabset-pill}

## Consumer Intentions


```{r descriptives consumer intentions, warning = FALSE, echo = FALSE}
CI_framing_table <- data %>%
  dplyr::group_by(framing_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(consumer_intentions, na.rm = TRUE),
            SD = sd(consumer_intentions, na.rm = TRUE))

knitr::kable(CI_framing_table, digits = 2, caption = "Consumer Intentions across Framing Conditions", col.names = c("Framing Condition", "n", "Mean", "SD"))
```

```{r warning = FALSE, echo = FALSE}
CI_norm_table <- data %>%
  dplyr::group_by(norm_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(consumer_intentions, na.rm = TRUE),
            SD = sd(consumer_intentions, na.rm = TRUE))

knitr::kable(CI_norm_table, digits = 2, caption = "Consumer Intentions across Norm Conditions", col.names = c("Norm Condition", "n", "Mean", "SD"))
```

## Consumer Behaviors

```{r label levels of consumer behaviors, echo=FALSE}
# 1 = I would like to enter into the raffle for a $50 gift card to spend on **new clothing items.
# 2 = I would like to enter into the raffle for a $50 gift card to spend on **secondhand clothing items.
levels(data$consumer_behaviors) <- c("New Clothing", "Secondhand Clothing")
```
```{r descriptives consumer behaviors, echo = FALSE}
CB_framing_table <- data %>%
  filter(!is.na(consumer_behaviors)) %>%
  dplyr::group_by(framing_condition, consumer_behaviors) %>% 
  dplyr::summarize(n = n()) %>%
  dplyr::mutate(proportion = n / sum(n))

knitr::kable(CB_framing_table, digits = 2, caption = "Proportion of Pro-Environmental Consumer Behaviors across Framing Conditions", col.names = c("Framing Condition", "Consumer Behaviors", "n", "Proportion"))
```


```{r echo = FALSE}
CB_norm_table <- data %>%
  filter(!is.na(consumer_behaviors)) %>%
  dplyr::group_by(norm_condition, consumer_behaviors) %>% 
  dplyr::summarize(n = n()) %>%
  dplyr::mutate(proportion = n / sum(n))

knitr::kable(CB_norm_table, digits = 2, caption = "Proportion of Pro-Environmental Consumer Behaviors across Norm Conditions", col.names = c("Framing Condition", "Consumer Behaviors", "n", "Proportion"))
```


## Values


```{r descriptives values, echo = FALSE}
valuesF_table <- data %>%
  dplyr::group_by(framing_condition) %>% 
  dplyr::summarize(n = n(),
            Mean_Bio = mean(biospheric, na.rm = TRUE),
            SD_Bio = sd(biospheric, na.rm = TRUE),
            Mean_Ego = mean(egoistic, na.rm = TRUE),
            SD_Ego = sd(egoistic, na.rm = TRUE))

knitr::kable(valuesF_table, digits = 2, caption = "Biospheric and Egoistic Values across Framing Conditions", col.names = c("Framing Condition", "n", "Mean Biospheric","SD","Mean Egoistic", "SD"))
```

```{r echo = FALSE}
valuesN_table <- data %>%
  dplyr::group_by(norm_condition) %>% 
  dplyr::summarize(n = n(),
            M_Biospheric = mean(biospheric, na.rm = TRUE),
            SD_Biospheric = sd(biospheric, na.rm = TRUE),
            M_Egoistic = mean(egoistic, na.rm = TRUE),
            SD_Egoistic = sd(egoistic, na.rm = TRUE))

knitr::kable(valuesN_table, digits = 2, caption = "Biospheric, Altruistic, Egoistic, and Hedonic Values across Norm Conditions", col.names = c("Norm Condition", "n", "Mean Biospheric","SD","Mean Egoistic", "SD"))

```


## Ingroup Identification

```{r descriptives ingroup ident, echo = FALSE}
ingroupF_table <- data %>%
  dplyr::group_by(framing_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(ingroup_identification, na.rm = TRUE),
            SD = sd(ingroup_identification, na.rm = TRUE))

knitr::kable(ingroupF_table, digits = 2, caption = "Ingroup Identification across Framing Conditions", col.names = c("Framing Condition", "n", "Mean", "SD"))
```

```{r echo = FALSE}
ingroupN_table <- data %>%
  dplyr::group_by(norm_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(ingroup_identification, na.rm = TRUE),
            SD = sd(ingroup_identification, na.rm = TRUE))

knitr::kable(ingroupN_table, digits = 2, caption = "Ingroup Identification across Norm Conditions", col.names = c("Norm Condition", "n", "Mean", "SD"))
```


## Clothing Interest

```{r descriptives clothing interest, echo = FALSE}
clothingF_table <- data %>%
  dplyr::group_by(framing_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(clothing_interest, na.rm = TRUE),
            SD = sd(clothing_interest, na.rm = TRUE))

knitr::kable(clothingF_table, digits = 2, caption = "Clothing Interest across Framing Conditions", col.names = c("Framing Condition", "n", "Mean", "SD"))
```

```{r echo = FALSE}
clothingN_table <- data %>%
  dplyr::group_by(norm_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(clothing_interest, na.rm = TRUE),
            SD = sd(clothing_interest, na.rm = TRUE))

knitr::kable(clothingN_table, digits = 2, caption = "Clothing Interest across Norm Conditions", col.names = c("Norm Condition", "n", "Mean", "SD"))

```

## Age

```{r echo = FALSE}
ageF_table <- data %>%
  dplyr::group_by(framing_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(Age, na.rm = TRUE),
            SD = sd(Age, na.rm = TRUE))

knitr::kable(ageF_table, digits = 2, caption = "Age across Framing Conditions", col.names = c("Framing Condition", "n", "Mean", "SD"))
```

```{r echo = FALSE}
ageN_table <- data %>%
  dplyr::group_by(norm_condition) %>% 
  dplyr::summarize(n = n(),
            Mean = mean(Age, na.rm = TRUE),
            SD = sd(Age, na.rm = TRUE))

knitr::kable(ageN_table, digits = 2, caption = "Age across Norm Conditions", col.names = c("Norm Condition", "n", "Mean", "SD"))
```


## Gender
```{r echo = FALSE}
genderF_table <- data %>%
  filter(!is.na(Gender)) %>%
  dplyr::group_by(framing_condition, Gender) %>% 
  dplyr::summarize(n = n()) %>%
  dplyr::mutate(proportion = n / sum(n))

knitr::kable(genderF_table, digits = 2, caption = "Proportion of Genders across Framing Conditions", col.names = c("Framing Condition", "Gender", "n", "Proportion"))
```

```{r echo = FALSE}
genderN_table <- data %>%
  filter(!is.na(Gender)) %>%
  dplyr::group_by(norm_condition, Gender) %>% 
  dplyr::summarize(n = n()) %>%
  dplyr::mutate(proportion = n / sum(n))

knitr::kable(genderN_table, digits = 2, caption = "Proportion of Genders across Norm Conditions", col.names = c("Norm Condition", "Gender", "n", "Proportion"))
```

## Notes

**Pro-Environmental Consumer Intentions:**

* Intentions to engage in pro-environmental consumer behaviors were similar across all framing conditions, but were descriptively highest when the decision was prefaced by a pro-environmental framing and lowest when the decision was prefaced by no framing (control condition). 
* Intentions to engage in pro-environmental consumer behaviors were also similar across all norm conditions, but were descriptively highest when the decision was prefaced by a description of a convention for consuming pro-environmentally and lowest when the decision was prefaced by a description of a social norm for consuming pro-environmentally.

**Pro-Environmental Consumer Behaviors:**

* When a pro-environmental consumer behavior was prefaced by the control condition (no framing) and by the self-enhancing framing, people opted for a gift card to spend on new clothing more than they opted for a gift card to spend on secondhand clothing. When a pro-environmental framing was used, however, the proportion of choosing a gift card to spend on secondhand clothing was higher than the proportion of people choosing a gift card to spend on new clothing.
* The distribution of people choosing the a gift card to spend on new versus on secondhand clothing was similar across all norm intervention conditions. There were slightly more than 50% of participants choosing the gift card to spend on new clothing across all norm conditions. 

**Covariates:**

* The distribution of participants' levels on all covariates (values, ingroup identification, clothing interest, age, gender) was similar across all levels of the framing and norm intervention conditions.


# Univariate Data Visualizations 

* Setting customized preferences for text on visualizations

```{r}
text_settings <- 
  theme(plot.title = element_text(size = 12, face = 'bold')) +
  theme(plot.subtitle = element_text(size = 8)) +
  theme(axis.title.x = element_text(size = 10, face = 'bold')) +
  theme(axis.title.y = element_text(size = 10, face = 'bold')) +
  theme(axis.text.x = element_text(size = 9)) +
  theme(axis.text.y = element_text(size = 10)) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```


## Consumer Intentions {.tabset .tabset-pill}

### Across Norm Conditions
```{r norm condition means viz, echo = FALSE}
# Consumer Intentions
ggplot(data = CI_norm_table, aes(x = norm_condition, fill = norm_condition, y = Mean)) +
  geom_bar(stat = "identity", color = 'black') +
  labs(title    = "Pro-Environmental Consumer Intentions across Norm Conditions",
       x        = "Norm Condition",
       y        = "Average Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Norm Condition") +
  theme_bw() +
  text_settings
```

* Pro-environmental consumer intentions were similar across all norm conditions but slightly higher in the convention condition than the other norm conditions. 

### Across Framing Conditions

```{r framing condition means viz, echo = FALSE}
# Consumer Intentions
ggplot(data = CI_framing_table, mapping = aes(x = framing_condition, fill = framing_condition, y = Mean)) +
  geom_bar(stat = "identity", color = 'black') +
  labs(title    = "Pro-Environmental Consumer Intentions across Framing Conditions",
       x        = "Framing Condition",
       y        = "Average Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Framing Condition") +
  theme_bw() +
  text_settings
```

* Pro-environmental consumer intentions were slightly higher in the pro-environmental framing condition compared to the other framing conditions.


## Consumer Behaviors {.tabset .tabset-pill}

### Across Norm Conditions

```{r echo = FALSE}
ggplot(data = CB_norm_table, aes(x = norm_condition, y = proportion, fill = consumer_behaviors)) +
  geom_bar(stat = "identity", color = 'black', position = "dodge") +
  labs(title = "Proportion Choosing New vs Secondhand Clothing",
       x = "Norm Condition",
       y = "Proportion") +
  scale_fill_discrete(name = "Consumer Behaviors") +
  theme_bw() +
  text_settings
```

* The proportion of people opting for a gift card to spend on new clothing (environmentally harmful decision) instead of on secondhand clothing (pro-environmental decision) was higher across all norm conditions.
  + However, the difference was smallest in the convention condition and largest when no normative message was used (control). 

### Across Framing Conditions
```{r echo = FALSE}
ggplot(data = CB_framing_table, aes(x = framing_condition, y = proportion, fill = consumer_behaviors)) +
  geom_bar(stat = "identity", color = 'black', position = "dodge") +
  labs(title = "Proportion Choosing New vs Secondhand Clothing",
       x = "Framing Condition",
       y = "Proportion") +
  scale_fill_discrete(name = "Consumer Behaviors") +
  theme_bw() +
  text_settings
```


* The proportion of people opting for a gift card to spend on secondhand clothing (pro-environmental decision) rather than on new clothing (environmentally harmful decision) was higher only when the choice was prefaced by a pro-environmental framing.
  + The proportion of people opting for a gift card to spend new clothing rather than secondhand clothing was highest when no framing was used (control).

## Values {.tabset .tabset-pill}

### Biospheric Values
```{r histograms of covariates, warning=FALSE, echo = FALSE}
# Values
bio_viz <- ggplot(data, aes(x = biospheric, y = ..density..)) +
  geom_histogram(fill = "forestgreen", color = "black", bins = 20) +
  geom_density(color = "blue", size = 1.1) +
  labs(x = "Biospheric Values",
       y = "Proportion of Sample") +
  theme_bw()

bio_viz
```

* The participants highly endorsed biospheric values. 
  + Scores on biospheric values were negatively skewed, suggesting our sample represents people high on this measure but not many people mid- to low on biospheric values. 

### Egoistic Values
```{r echo = FALSE}
ego_viz <- ggplot(data, aes(x = egoistic, y = ..density..)) +
  geom_histogram(fill = "darkviolet", color = "black", bins = 20) +
  geom_density(color = "blue", size = 1.1) +
  labs(x = "Egoistic Values",
       y = "Proportion of Sample") +
  theme_bw()

ego_viz
```

* Participants' scores on egoistic values were approximately normally distributed.
  + The mean was above the midpoint of the scale (*M* = 4.99, *SD* = 0.92), suggesting the sample on average leaned towards endorsing egoistic values. 

## Ingroup Identification
```{r echo = FALSE}
ingroup_viz <- ggplot(data, aes(x = ingroup_identification, y = ..density..)) +
  geom_histogram(fill = "deeppink3", color = "black", bins = 20) +
  geom_density(color = "blue", size = 1.1) +
  labs(x = "Ingroup Identification",
       y = "Proportion of Sample") +
  theme_bw()

ingroup_viz
```

* Participants' scores on ingroup identification were approximately normally distributed.
  + The mean was slightly above the midpoint of the scale (*M* = 4.64, *SD* = 1.01), suggesting participants tended identify with the reference group (other University of Oregon students).

## Interest in Clothing
```{r echo = FALSE}
cloth_viz <- ggplot(data, aes(x = clothing_interest, y = ..density..)) +
  geom_histogram(fill = "indianred2", color = "black", bins = 20) +
  geom_density(color = "blue", size = 1.1) +
  labs(x = "Interest in Clothing",
       y = "Proportion of Sample") +
  theme_bw()

cloth_viz
```

* Participants' scores on interest in clothing were approximately normally distributed. 
  + The mean was near the midpoint of the scale (*M* = 3.13, *SD* = 0.80), suggesting about half of the sample expressed interest in clothing and about half did not express interest in clothing. 

# Bivariate Data Visualizations 

## Biospheric Values and Consumer Intentions {.tabset .tabset-pill}

### Overall
```{r echo = FALSE}
ggplot(data = data, aes(x = biospheric, y = consumer_intentions)) +
  geom_point(color = "purple", alpha = 0.8) +
  geom_smooth(method = "lm", color = "deepskyblue2") +
  labs(title    = "Biospheric Values and Pro-environmental Consumer Intentions",
       x        = "Biospheric Values",
       y        = "Pro-Environmental Consumer Intentions") +
  theme_bw() +
  text_settings
```

### Across Norm Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = biospheric, y = consumer_intentions, color = norm_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Biospheric Values and Pro-environmental Consumer Intentions",
       x        = "Biospheric Values",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Norm Condition") +
  facet_wrap(~norm_condition) +
  theme_bw() +
  text_settings
```

### Across Framing Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = biospheric, y = consumer_intentions, color = framing_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Biospheric Values and Pro-environmental Consumer Intentions",
       x        = "Biospheric Values",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Framing Condition") +
  facet_wrap(~framing_condition) +
  theme_bw() +
  text_settings
```

### Notes

* There was a general positive association between endorsement of biospheric values and intentions to engage in pro-environmental consumer behaviors. 
  + This relationship was consistent across the framing and norm intervention conditions. 

## Egoistic Values and Consumer Intentions {.tabset .tabset-pill}

### Overall
```{r echo = FALSE}
ggplot(data = data, aes(x = egoistic, y = consumer_intentions)) +
  geom_point(color = "darkorange", alpha = 0.8) +
  geom_smooth(method = "lm", color = "deepskyblue2") +
  labs(title    = "Egoistic Values and Pro-environmental Consumer Intentions",
       x        = "Egoistic Values",
       y        = "Pro-Environmental Consumer Intentions") +
  theme_bw() +
  text_settings
```

### Across Norm Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = egoistic, y = consumer_intentions, color = norm_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Egoistic Values and Pro-environmental Consumer Intentions",
       x        = "Egoistic Values",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Norm Condition") +
  facet_wrap(~norm_condition) +
  theme_bw() +
  text_settings
```

### Across Framing Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = egoistic, y = consumer_intentions, color = framing_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Egoistic Values and Pro-environmental Consumer Intentions",
       x        = "Egoistic Values",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Framing Condition") +
  facet_wrap(~framing_condition) +
  theme_bw() +
  text_settings
```

### Notes

* There was a general negative association between endorsement of egoistic values and intentions to engage in pro-environmental consumer behaviors. 
  + This relationship was consistent across the framing and norm intervention conditions. 

## Interest in Clothing and Consumer Intentions {.tabset .tabset-pill}

### Overall
```{r echo = FALSE}
ggplot(data = data, aes(x = clothing_interest, y = consumer_intentions)) +
  geom_point(color = "deeppink3", alpha = 0.8) +
  geom_smooth(method = "lm", color = "deepskyblue2") +
  labs(title    = "Interest in Clothing and Pro-environmental Consumer Intentions",
       x        = "Interest in Clothing",
       y        = "Pro-Environmental Consumer Intentions") +
  theme_bw() +
  text_settings
```

### Across Norm Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = clothing_interest, y = consumer_intentions, color = norm_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Interest in Clothing and Pro-environmental Consumer Intentions",
       x        = "Interest in Clothing",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Norm Condition") +
  facet_wrap(~norm_condition) +
  theme_bw() +
  text_settings
```

### Across Framing Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = clothing_interest, y = consumer_intentions, color = framing_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Interest in Clothing and Pro-environmental Consumer Intentions",
       x        = "Interest in Clothing",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Framing Condition") +
  facet_wrap(~framing_condition) +
  theme_bw() +
  text_settings
```

### Notes

* There was little to no association between participants' interest in clothing and their intentions to engage in pro-environmental consumer behaviors. 
  + This lack of a relationship was generally consistent across the framing and norm intervention conditions. 


## Ingroup Identification and Consumer Intentions {.tabset .tabset-pill}

### Overall
```{r echo = FALSE}
ggplot(data = data, aes(x = ingroup_identification, y = consumer_intentions)) +
  geom_point(color = "darkred", alpha = 0.8) +
  geom_smooth(method = "lm", color = "deepskyblue2") +
  labs(title    = "Ingroup Identification and Pro-environmental Consumer Intentions",
       x        = "Ingroup Identification",
       y        = "Pro-Environmental Consumer Intentions") +
  theme_bw() +
  text_settings
```

### Across Norm Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = ingroup_identification, y = consumer_intentions, color = norm_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Ingroup Identification and Pro-environmental Consumer Intentions",
       x        = "Ingroup Identification",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Norm Condition") +
  facet_wrap(~norm_condition) +
  theme_bw() +
  text_settings
```

### Across Framing Conditions
```{r echo = FALSE}
ggplot(data = data, aes(x = ingroup_identification, y = consumer_intentions, color = framing_condition)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm") +
  labs(title    = "Ingroup Identification and Pro-environmental Consumer Intentions",
       x        = "Ingroup Identification",
       y        = "Pro-Environmental Consumer Intentions") +
  scale_fill_discrete(name = "Framing Condition") +
  facet_wrap(~framing_condition) +
  theme_bw() +
  text_settings
```

### Notes

* There was a slightly positive, but not substantial, association between the degree to which participants identified with the reference group (other UO students) and their intentions to engage in pro-environmental consumer behaviors. 
